The Milky Way Project: Leveraging Citizen Science and Machine Learning to Detect Interstellar Bubbles

### 2013-10-03 paper/ms.tex
files at time of submission

\title{Automated Detection of Interstellar Bubbles}
\shortauthors{Beaumont et al}
\author{Christopher N. Beaumont$^{1,2}$, Alyssa A. Goodman$^2$, Sarah Kendrew$^3$, Jonathan P. Williams$^1$, Robert Simpson$^3$}
\affil{$^1$Institute for Astronomy, University of Hawai'i, 2680 Woodlawn Drive, Honolulu HI 96822;  beaumont@ifa.hawaii.edu}
\affil{$^2$Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge MA 02138}
\affil{$^3$Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK}

\begin{abstract}
We present Brut, an algorithm to identify bubbles in infrared images of the Galactic midplane. Brut is based on the Random Forest algorithm, and uses bubbles identified by citizen scientists from the Milky Way Project to discover the identifying characteristics of bubbles in images from the \textit{Spitzer} Space Telescope. We demonstrate that Brut's ability to identify bubbles is comparable to expert astronomers. We use Brut to re-assess the bubbles in the Milky Way Project catalog, and find that $10-30\%$ of the objects in the Milky Way Project catalog are non-bubble interlopers. Relative to these interlopers, high-reliability bubbles are more confined to the mid plane, and display a stronger excess of Young Stellar Objects near bubble rims. Furthermore, Brut is able to discover bubbles missed by previous searches -- particularly bubbles near bright sources which have low contrast relative to their surroundings. Brut demonstrates the synergies that exist between citizen scientists, professional scientists, and machine learning techniques.
\end{abstract}

\section{Introduction}
\label{sec:intro}
Stellar feedback has an important influence on the dynamics and energy balance of the Interstellar Medium (ISM) \citep{Zinnecker07}. Winds and radiation fields from massive young stars can reshape nearby molecular clouds. This interaction can replenish energy dissipated by turbulence, trigger star formation by compressing and collecting gas, and even chemically dissociate or physically disperse molecular clouds.

Interstellar ``bubbles'' are a primary mode of stellar feedback. Young OB stars have sufficient stellar winds and ionizing photon luminosity to sculpt spherical or ring-like cavities in their surrounding molecular clouds. Bubbles in particular are relevant because, compared to collimated outflows, they affect a larger volume of ambient molecular clouds and, compared to supernovae, occur around a larger proportion of stars.

Unfortunately, due to their complex morphologies, bubbles -- like many features of the interstellar medium -- are difficult to identify and analyze. Existing catalogs of spatially extended bubbles have typically been identified manually \citep{Churchwell06, Churchwell07, Simpson12, Helfand06}. This has two main disadvantages. First, it is cumbersome and increasingly infeasible as datasets grow ever larger. Second, manual classification is inherently subjective and non-repeatable; humans are susceptible to fatigue, boredom, and subtle selection biases whose impact on the resulting catalog is difficult to calibrate. The problems associated manual bubble detection are germane to many analyses with a subjective component.

@@

Our goal in this work is to apply machine learning techniques to the task of bubble detection, to evaluate the potential of this approach. Using a catalog of known bubbles identified by the citizen scientists of the Milky Way Project \citep{Simpson12}, we ``teach'' an algorithm to identify bubbles in image data from the \textit{Spitzer} space telescope. We describe the design of this algorithm, which we call Brut, in Section \ref{sec:method}. In Section \ref{sec:expert}, we use a set of expert classifications to measure Brut's performance at bubble detection. In Section \ref{sec:prob}, we demonstrate that this detector produces useful probabilistic estimates for whether any particular image contains a bubble -- these probabilities correlate well with how astronomers classify similar regions. We use this detector to look for biases and incompleteness in existing bubble catalogs. This analysis yields a new catalog of high-probablity bubbles, and we explore how the ensemble properties of this catalog differ from old catalog. In In Section \ref{sec:blind}, we apply Brut to the task of discovering bubbles missing from current catalogs. Section \ref{sec:next_steps}, we consider how this approach applies more generally to future data analysis problems in astronomy.

@@

Generally speaking, a ``bubble'' is a shell-like, 1-30 parsec-scale cavity in the ISM, cleared by a combination of thermal overpressure, radiation pressure, and stellar winds. The basic structure of a bubble is shown in Figure \ref{fig:schematic}. \cite{Stromgren39} first worked out the ionization structure of OBA stars in the ISM. The act of ionizing hydrogen heats gas from 100 to 10$^4$K, creating a strong overpressure that drives expansion into the surrounding medium. This expansion creates an overdense shell of gas along the expansion front. \cite{Castor75} and \cite{Weaver77} extended this analysis to include the effects of strong stellar winds from O and early B stars. The main effect of a stellar wind is to create a shocked, high-temperature (10$^6$K) region within the 10$^4$K ionization region. This shock sweeps up additional material within the ionization region, potentially creating a second shell. Though the term ``bubble'' originally referred to cavities cleared by stellar winds, modern observational studies of ``bubbles'' tend to include both these objects and classical \hii\, regions. Throughout this paper, we use the term ``bubble'' more widely to refer to any shell-like cavity cleared by a young star.

The Spitzer Space Telescope and its surveys of the Galactic midplane -- {\sc glimpse} \citep{Benjamin03} and {\sc mipsgal} \citep{Carey09} -- enabled comprehensive, statistical studies of bubbles in the Galaxy. Mid-infrared wavelengths are well-suited for bubble observations, as they penetrate deeper into the Galactic disk and match bubble emission features. Figure \ref{fig:schematic} schematically depicts the observational signature of a bubble in Spitzer data. The interface between bubble shells and the ambient ISM excites polycyclic aromatic hydrocarbon (PAH) emission features, several of which overlap the 8\um\, bandpass. Bubble interiors often emit at 24\um, due to emission from hot dust grains \citep{Everett10}.

\cite{Churchwell06} and \cite{Churchwell07} carried out the first search for Bubbles in Spitzer images of the Galactic plane, yielding  a catalog of some 600 bubbles located at $|\ell| < 65^\circ$, $|b| < 1^\circ$. These objects were identified by four astronomers manually searching through 3.5-8.0 \um\, images (they did not have access to 24 \um\, images). They noted that each astronomer possessed different selection biases, and cautioned that this catalog was likely incomplete.

In an attempt to overcome the inherent bias in manual classification, the web-based Milky Way Project \citep{Simpson12} enlisted over 35,000 citizen scientists to search for bubbles in Spitzer data. The Milky Way Project (hereafter MWP) presented pre-scaled Spitzer images at 4.5\um,  8\um, and 24\um, and asked citizen scientists to draw ellipses around potential bubbles. Figure \ref{fig:mwp} shows two typical images presented to the public. This led to a dramatically larger catalog of $\sim 5000$ objects. The organizers of the MWP attribute this large increase to the 1000-fold increase in human classifiers and the use of 24\um\, data to further emphasize bubble interiors. They estimate the MWP catalog to be 95\% complete, based on the falloff rate at which new objects were discovered. Still, they cautioned that this catalog is heterogenous, and probably affected by hard-to-measure selection biases.

\begin{figure}[h!]
\includegraphics[width=3in]{schematic}
\caption{Schematic structure of a bubble, adapted from \citep{Freyer03}. The red and green colors encode the typical morphology seen in Spitzer images, where green is assigned to 8 \um\, emission -- dominated by PAH fluorescence -- and red is assigned to 24 \um emission -- dominated by hot dust.}
\label{fig:schematic}
\end{figure}

@@

\begin{figure*}
\includegraphics[trim= .7in 0 0 0, clip]{gallery}
\caption{Different astrophysical objects in the MWP catalog. a) ``Canonical'' wind-blown bubbles and \hii\, regions. b) shells
without 8 \um\, PAH emission (likely supernovae and planetary nebulae). c) generic ISM structures of unclear astrophysical origin.}
\label{fig:gallery}
\end{figure*}

@@

First, human classification is time consuming, and people-hours are a limited resource. Even by enlisting large communities of citizen scientists, data from next generation surveys will be too large to search exhaustively. For example, the $>$ 35,000 citizen scientists of the Milky Way Project classified roughly 45 GB of image data from Spitzer. Current and next-generation astronomical datasets are many thousands of times larger than this, suggesting tens of millions of citizen scientists would be needed for similar exhaustive, manual searches.

Second, many scientifically important tasks are not suitable for enlisting the public. Part of the success of the Milky Way Project is due to the fact that the Spitzer images are beautiful, contain many bubbles, and are compelling to browse through. Searches for very rare or less aesthetic objects, on the other hand, are more tedious and less likely to entice large citizen science communities.

Finally, manual classification is not easily repeatable, and hard to calibrate statistically. For example, it is unknown how well the consensus opinion among citizen scientists corresponds to consensus among astronomers. The MWP catalog does not include any estimate of the probability that each object is a real bubble, as opposed to other structures in the ISM.

Automatic classifications driven by machine learning techniques nicely complement human classification. Such an approach easily scales to large data volumes and is immune to some of the factors that affect humans, like boredom and fatigue. Furthermore, because algorithmic classifications are systematic and repeatable, they are easier to interpret and statistically characterize.

@@

\subsection{Training Data}
\label{sec:method_training_data}
The MWP catalog is a natural choice for a training dataset. However, as shown in Figure \ref{fig:gallery}, the catalog as a whole contains many examples of objects that are not bubbles, and can compromise the training process. Thus, we manually curated a list of 468 objects in the Milky Way Project catalog of large objects which were clear examples of bubbles. We focus our attention on objects in the large bubble catalog since the web interface used to build the small catalog was different, and may have different selection biases \citep{Simpson12}. The large bubble catalog contains 3744 out of 5106 total bubbles in the MWP catalog, all of which have semimajor axes $a \geq 13.5''$.

We also require an example set of non-bubbles. In principle, any random region of the sky that doesn't overlap an object in the MWP catalog is a candidate negative example -- while some of these fields actually contain bubbles not in the MWP catalog, they are rare enough that they don't  impact the training. However, we found that selecting negative examples at random was sub-optimal. The reason for this is that most random fields are nearly empty, and easily distinguished from fields with bubbles. Classifiers generated from such a dataset incorrectly labeled nearly any field containing ISM structure as a bubble.

A better set of negative examples includes more examples of fields containing structure (Figure \ref{fig:bootstrap_neg}). We built such a collection in a bootstrapped fashion. Starting with random negative examples, we trained a classifier and used it to scan a large number (20,000) of negative fields. We then discarded half of the initial negative examples (those classified most confidently as not containing bubbles), and replaced them with a random sample of the mis-classified examples. We repeated this process several times, but found that one iteration was usually sufficient to build a good set of training data.

\begin{figure}[h!]
\includegraphics{bootstrap_neg}
\caption{Examples of typical randomly selected negative examples (left), and ``hard'' examples (right)}
\label{fig:bootstrap_neg}
\end{figure}

@@

Our feature vectors have been designed based on insights in the automated face detection literature \citep{ViolaJones}.  The basic strategy is to encode a very large number ($\sim$40,000) of generic statistics about extended structures at various positions and scales. This strategy seems to be more effective then trying to tune a feature vector to the specific object being identified. While most elements in the feature vector will have no predictive power in the classification, the appropriate learning algorithm will simply ignore this information.

@@

\begin{enumerate}
\item The wavelet coefficients from the Discrete Cosine and Daub4 wavelet transforms. These coefficients encode how much power is in various spatial frequencies.
\item The image dot product\footnote{i.e., the sum of the pixel-by-pixel product of two images} of each image with template images of rings of different size and thickness. Bubbles are morphologically similar to these templates, and presumably have larger dot products than a random intensity distribution.
\item The byte size of each compressed postage stamp image -- images substantially more or less smooth than a bubble compress to smaller and larger files, respectively.
\item The DAISY features of the image \citep{DAISY}, which characterize the local gradients in an image and have been useful in other feature recognition contexts.
\end{enumerate}

\subsection{Training}
While there are many algorithms designed to learn class labels form a training set of feature vectors, we focus here on the Random Forest algorithm \citep{Breiman01}. Random Forests are aggregates of a simpler learning algorithm called a decision tree. A decision tree makes classifications by passing each feature vector through the nodes of a tree diagram. Each node consists of a simple boolean test of one element of the feature vector (e.g., $F_3 > 5$). Feature vectors traverse the left or right paths through the tree based on whether each test is true or false, until they reach a terminal node. The terminal node contains the final classification (in the present context, ``Bubble'' or ``Not Bubble''). Decision trees are constructed in a greedy, stage-wise fashion: each boolean test is chosen to separate the set of training data into two subclasses that are more cleanly partitioned between classes. Several different heuristics can be used to measure how ``clean'' the partition is.

On their own, decision trees are prone to over-fitting the training data by adding too many nodes. Random Forests overcome this problem by building several trees on different subsets of the data. The final classification is made by taking the majority vote of each subtree. This mitigates over-fitting in the individual trees, since over-fitting is sensitive to the specific subset of data used, and hence averages away. Random forests have proven effective in many machine learning contexts \citep{Kuhn13}.

In addition to their good performance and robustness against over fitting, Random Forests have several other advantageous properties: it is possible to interpret how important each element in the feature vector is to the classification, the algorithm naturally ignores irrelevant elements of the feature vector, and it is conceptually and computationally simple.

\subsection{Hyperparameter Optimization}
The Random Forest algorithm has a small number of tunable parameters. The most important of these are the number of trees in the forest, and the choice of splitting heuristic used when building each tree. We determined good settings for these hyper-parameters using cross validation. We partitioned the training data into two sets, and used the first set to train the algorithm with a given set of hyper-parameters. The second set of cross-validation data was used to evaluate the model. Then, the hyper parameters were adjusted and the process repeated. We chose the hyper-parameters that maximized the performance on the cross-validation data -- we chose forests with 800 trees, using the information gain splitting heuristic \citep{Raileanu04}.

@@

Classifiers cannot be used to re-classify the data used to train them, due to the possibility of over-fitting. Thus, to obtain valid classifications for all fields, we constructed three classifiers. Each classifier is trained using data from two-thirds of the sky, and used to classify the remaining third. The sky is partitioned into thirds based on the modulus of the nearest whole longitude: for example, the first classifier is trained on examples where $0.5 < (\ell \mod{3})  < 2.5$, and used to classify fields where ($\ell \mod{3}) < 0.5 ~ || ~(\ell \mod{3}) > 2.5 $

@@

One standard way to summarize the performance of a classifier is to plot the false positive rate (fraction of negative examples incorrectly classified) versus the true positive rate (fraction of bubbles currently classified) as the threshold varies. This is called the Receiver Operating Characteristic, and is shown for the three classifiers in Figure \ref{fig:roc}. The false positive rate is measured by classifying $\sim 15000$ random negative fields, and the true positive rate is measured using the high-quality subset of the MWP catalog described above. Recall that the set of bubbles in this test set were selected because they are very well defined. Thus, Figure \ref{fig:roc} overestimates the true positive rate when presented with more ambiguous bubbles. We compare the performance of the classifier on more representative bubbles in Section \ref{sec:expert}.

\begin{figure}[h!]
\includegraphics{roc}
\caption{The tradeoff between false positive rate and true positive rate, for each of the three classifiers in Brut.}
\label{fig:roc}
\end{figure}

@@

Using our automatic bubble classifier to conduct a full search for Bubbles in \textit{Spitzer} data is straightforward. The main task involves scanning the classifier over all images, at all scales. At each of these regions, the classifier yields a score, representing its confidence that the field contains a bubble. A simple threshold on this score identifies regions likely to contain a bubble -- from the discussion of Figure \ref{fig:roc} above, adjusting this threshold sets the tradeoff between completeness and reliability.

In practice, the classifier is insensitive to small adjustments to the size or location of the field. As a result, bubbles are usually associated with a cluster of postage stamps that rise above the detection threshold (the grey boxes in Figure \ref{fig:cluster}). We follow a simple, greedy procedure to merge these individual detections. Given a collection of postage stamp descriptions (location and scale), this procedure identifies the two nearest locations at similar scales, and discards the location assigned a lower score by the classifier. This process repeats until no two fields are close enough to discard. Two fields are close enough to merge if they are separated by less than the radius of either field, and have sizes that differ by less than 50\%. The blue box in Figure \ref{fig:cluster} shows the result of merging detections.

\begin{figure}[h!]
\includegraphics{cluster}
\caption{An illustration of how Brut merges multiple bubble detections.}
\label{fig:cluster}
\end{figure}

@@

\begin{comment}
\subsection{Classifying the l=35 field}
\label{sec:case_study}
To better understand the successes and failures of our classifier, we have analyzed the longitude range $34.5^\circ < \ell < 35.5^\circ$ in detail. In our first analysis, we scanned for all bubbles above a minimum size $r _{\rm min} = 80''$ \todo{Is this r or d?}, with scores above $s_{\rm min} = 0.55$. This search yielded 25 bubbles after merging.

Among the 25 automatic detections, 13 are unambiguous bubbles with clear counterparts in the MWP catalog. A further four objects appear to be bubbles as judged by the authors, but are part of bubble complexes. While the complexes are identified at least once in both catalogs, each catalog decomposes the complex slightly differently. Thus, these 4 objects don't have an obvious one-to-one relationship with an object in the MWP catalog. Figure \ref{fig:cluster_confusion} shows an example of one region of confusion. The automated classifier discovers two new candidate bubbles with no counterpart in the MWP catalog. \todo{Lookup what the MWP images for these regions look like. Why weren't they detected? My guess is contrast}. The remaining 6 objects in the automated catalog are false positives -- three are regions containing a bright and/or extended 24\um\, point source confused for a bubble interior, and three are regions with 8 \um\, nebulosity that don't appear shell-like.

>>>>

\title{The Milky Way Project: Leveraging Citizen Science and Machine Learning to Detect Interstellar Bubbles}

\shortauthors{Beaumont et al}
\author{Christopher N. Beaumont$^{1,2}$, Alyssa A. Goodman$^2$, Sarah Kendrew$^3$, Jonathan P. Williams$^1$, Robert Simpson$^3$}
\affil{$^1$Institute for Astronomy, University of Hawai'i, 2680 Woodlawn Drive, Honolulu HI 96822;  beaumont@ifa.hawaii.edu}
\affil{$^2$Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge MA 02138}
\affil{$^3$Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK}

\begin{abstract}
We present Brut, an algorithm to identify bubbles in infrared images of the Galactic midplane. Brut is based on the Random Forest algorithm, and uses bubbles identified by $>35,000$ citizen scientists from the Milky Way Project to discover the identifying characteristics of bubbles in images from the \textit{Spitzer Space Telescope}. We demonstrate that Brut's ability to identify bubbles is comparable to expert astronomers. We use Brut to re-assess the bubbles in the Milky Way Project catalog, and find that $10-30\%$ of the objects in this catalog are non-bubble interlopers. Relative to these interlopers, high-reliability bubbles are more confined to the mid plane, and display a stronger excess of Young Stellar Objects along and within bubble rims. Furthermore, Brut is able to discover bubbles missed by previous searches -- particularly bubbles near bright sources which have low contrast relative to their surroundings. Brut demonstrates the synergies that exist between citizen scientists, professional scientists, and machine learning techniques. In cases where ``untrained'' citizens can identify patterns that machines cannot detect without training, machine learning algorithms like Brut can use the \emph{output} of citizen science projects as \emph{input} training sets, offering tremendous opportunities to speed the pace of scientific discovery.
\end{abstract}

\section{Introduction}
\label{sec:intro}
Stellar feedback has an important influence on the dynamics and energy balance of the Interstellar Medium (ISM) \citep{Zinnecker07}. Winds and radiation fields from massive young stars can reshape nearby molecular clouds. This interaction can replenish energy dissipated by turbulence, trigger star formation by compressing and collecting gas, and even chemically dissociate or physically disperse molecular clouds \citep{Matzner02}.

Interstellar ``bubbles'' are a primary manifestation of stellar feedback. Young OB stars have sufficient stellar winds and ionizing photon luminosity to sculpt spherical or ring-like cavities in their surrounding molecular clouds. Bubbles in particular are relevant because, compared to collimated outflows, they affect a larger volume of ambient molecular clouds and, compared to supernovae, occur around a larger proportion of stars and persist for a longer period of time \citep{Matzner02, Arce11}.

Unfortunately, due to their complex morphologies, bubbles -- like many features of the interstellar medium -- are difficult to identify and analyze. Existing catalogs of spatially extended bubbles have typically been identified visually \citep{Churchwell06, Churchwell07, Simpson12, Helfand06}. This has two main disadvantages. First, it is cumbersome and increasingly infeasible as datasets grow ever larger. Second, manual classification is inherently subjective and non-repeatable; humans are susceptible to fatigue, boredom, and subtle selection biases whose impact on the resulting catalog is difficult to calibrate. The problems associated manual bubble detection are germane to many analyses with a subjective component.

@@

Our goal in this work is to apply machine learning techniques to the task of bubble detection, and to evaluate the potential of this approach. Using a catalog of known bubbles identified by the citizen scientists of the Milky Way Project \citep{Simpson12}, we ``teach'' an algorithm to identify bubbles in image data from the \textit{Spitzer} space telescope. We describe the design of this algorithm, which we call Brut, in Section \ref{sec:method}. In Section \ref{sec:expert}, we use a set of expert classifications to measure Brut's performance at bubble detection. In Section \ref{sec:prob}, we demonstrate that this detector produces useful probabilistic estimates for whether any particular image contains a bubble -- these probabilities correlate well with how astronomers classify similar regions. We use this detector to look for biases and incompleteness in existing bubble catalogs. This analysis yields a new catalog of high-probability bubbles, and we explore how the ensemble properties of this catalog differ from the old catalog. In Section \ref{sec:blind}, we apply Brut to the task of discovering bubbles missing from current catalogs. In Section \ref{sec:next_steps}, we consider how this approach applies more generally to future data analysis problems in astronomy.

@@

Generally speaking, a bubble is a shell-like, 1-30 parsec-scale cavity in the ISM, cleared by a combination of thermal overpressure, radiation pressure, and stellar winds. The basic structure of a bubble is shown in Figure \ref{fig:schematic}. \cite{Stromgren39} first derived the ionization structure around OBA stars in the ISM. The act of ionizing hydrogen heats gas from 100 to 10$^4$K, creating a strong overpressure that drives expansion into the surrounding medium. This expansion creates an overdense shell of gas along the expansion front. \cite{Castor75} and \cite{Weaver77} extended this analysis to include the effects of strong stellar winds from O and early B stars. The main effect of a stellar wind is to create a shocked, high-temperature (10$^6$K) region within the 10$^4$K ionization region. This shock sweeps up additional material within the ionization region, potentially creating a second shell. Though the term ``bubble'' originally referred to cavities cleared by stellar winds, modern observational studies of ``bubbles'' tend to include both these objects and classical \hii\, regions. Furthermore, \cite{Beaumont10} demonstrated that many bubbles are embedded in relatively thin clouds, and more closely resemble rings than spheres. Throughout this paper, we use the term ``bubble'' broadly to refer to any ring- or shell-like cavity cleared by a young or main sequence star.

The \emph{Spitzer Space Telescope} and its surveys of the Galactic midplane -- {\sc glimpse} \citep{Benjamin03} and {\sc mipsgal} \citep{Carey09} -- enabled comprehensive, statistical studies of bubbles in the Galaxy. Mid-infrared wavelengths are well-suited for bubble observations, as they penetrate deeper into the Galactic disk and match bubble emission features. Figure \ref{fig:schematic} schematically depicts the observational signature of a bubble in \emph{Spitzer} data. The interface between bubble shells and the ambient ISM excites polycyclic aromatic hydrocarbon (PAH) emission features, several of which fall within \emph{Spitzer}'s 8\um\, bandpass. Bubble interiors often emit at 24\um, due to emission from hot dust grains \citep{Everett10}.

\cite{Churchwell06} and \cite{Churchwell07} carried out the first search for Bubbles in \emph{Spitzer} images of the Galactic plane, yielding  a catalog of some 600 bubbles located at $|\ell| < 65^\circ$, $|b| < 1^\circ$. These objects were identified by four astronomers manually searching through 3.5-8.0 \um\, images (they did not have access to 24 \um\, images). \cite{Churchwell06} noted that each astronomer possessed different selection biases, and cautioned that their catalog was likely incomplete.

In an attempt to overcome the inherent bias in manual classification, the web-based Milky Way Project \cite{Simpson12} enlisted over 35,000 citizen scientists to search for bubbles in \emph{Spitzer} data. The Milky Way Project (hereafter MWP) presented color-composite \emph{Spitzer} images at 4.5\um,  8\um, and 24\um, and asked citizen scientists to draw ellipses around potential bubbles. Figure \ref{fig:mwp} shows two typical images presented to the public. Each image in the Milky Way Project was 800x400 pixels, with a pixel scale ranging between 1.35$''$ and 6.75$''$ per pixel. The citizen science effort produced a dramatically larger catalog of $\sim 5000$ objects, nearly 10 times the number in the catalog of $\sim 600$ shells cataloged by the four astronomers of the \cite{Churchwell06, Churchwell07} surveys. The organizers of the MWP attribute this large increase to the 10,000-fold increase in human classifiers and the use of 24\um\, data to further emphasize bubble interiors. They estimate the MWP catalog to be 95\% complete, based on the falloff rate at which new objects were discovered. Still, they cautioned that this catalog is heterogenous, and probably affected by hard-to-measure selection biases.

\begin{figure}[h!]
\includegraphics[width=3in]{schematic}
\caption{Schematic structure of a bubble, adapted from \citep{Freyer03}. The red and green colors encode the typical morphology seen in \emph{Spitzer} images, where green is assigned to 8 \um\, emission -- dominated by PAH fluorescence -- and red is assigned to 24 \um\, emission -- dominated by hot dust.}
\label{fig:schematic}
\end{figure}

@@

\begin{figure*}
\includegraphics[trim= .7in 0 0 0, clip]{gallery}
\caption{Different astrophysical objects in the MWP catalog. a) ``Canonical'' wind-blown bubbles and \hii\, regions. b) shells
without 8 \um\, PAH emission (likely supernovae or bubbles around evolved massive stars). c) generic ISM structures of unclear astrophysical origin.}
\label{fig:gallery}
\end{figure*}

@@

First, human classification is time consuming, and people-hours are a limited resource. Even by enlisting large communities of citizen scientists, data from next generation surveys will be too large to search exhaustively. For example, the $>$ 35,000 citizen scientists of the Milky Way Project classified roughly 45 GB of image data from \emph{Spitzer}. Current and next-generation astronomical datasets are many thousands of times larger than this, suggesting tens of millions of citizen scientists would be needed for similar exhaustive searches through tera- and petabyte datasets.

Second, many scientifically important tasks are not suitable for enlisting the public. Part of the appeal of the Milky Way Project is due to the fact that the \emph{Spitzer} images are beautiful, contain many bubbles, and are compelling to browse through. Searches for very rare objects, or tasks where the scientific justification is less apparent to a citizen scientist, may be less likely to entice large volunteer communities. \cite{Raddick13} considers the motivations of citizen scientists in greater detail.

Finally, manual classification is not easily repeatable, and hard to calibrate statistically. For example, it is unknown how well the consensus opinion among citizen scientists corresponds to consensus among astronomers. The MWP catalog does not include any estimate of the probability that each object is a real bubble, as opposed to another structure in the ISM.

Automatic classifications driven by machine learning techniques nicely complement human classification. Such an approach easily scales to large data volumes and is immune to some of the factors that affect humans, like boredom and fatigue. Furthermore, because algorithmic classifications are systematic and repeatable, they are easier to interpret and statistically characterize. Despite the structural complexity of the ISM, \cite{Beaumont11} demonstrated that automatic classification algorithms can discriminate between different ISM structures based upon morphology.

@@

\subsection{Training Data}
\label{sec:method_training_data}
The MWP catalog is a natural choice for a training dataset. However, as shown in Figure \ref{fig:gallery}, the catalog as a whole contains many examples of objects that are not bubbles, and can compromise the training process. Thus, we manually curated a list of 468 objects in the Milky Way Project catalog which were clear examples of bubbles. We focus our attention on objects in the large bubble catalog, which contains 3744 out of 5106 total bubbles, with simimajor axes  $a \geq 13.5''$  \citep{Simpson12}. We focus on the large bubble catalog since the web interface used to build the small catalog was different, and may have different selection biases. The largest bubble in the MWP catalog has a semi major axis $a=10.3'$. This upper limit is set by the widest-field images presented to citizen scientists -- we discuss an implication of this cutoff in Section \ref{sec:prob}.

Training also requires an example set of non-bubbles. In principle, any random region of the sky that doesn't overlap an object in the MWP catalog is a candidate negative example -- while some of these fields actually contain bubbles not in the MWP catalog, they are rare enough that they don't  impact the training. However, we found that selecting negative examples at random was sub-optimal. The reason for this is that most random fields are nearly empty, and easily distinguished from fields with bubbles. Classifiers generated from such a dataset incorrectly labeled nearly any field containing ISM structure as a bubble.

A better set of negative examples includes more fields containing structure (Figure \ref{fig:bootstrap_neg}). We built such a collection in a bootstrapped fashion. Starting with random negative examples, we trained a classifier and used it to scan a large number (20,000) of negative fields. We then discarded half of the initial negative examples (those classified most confidently as not containing bubbles), and replaced them with a random sample of the mis-classified examples. We repeated this process several times, but found that one iteration was usually sufficient to build a good set of training data.

\begin{figure}[h!]
\includegraphics{bootstrap_neg}
\caption{Examples of typical randomly selected negative examples (left), and ``hard'' examples (right). Including more hard examples produces more discriminating classifiers.}
\label{fig:bootstrap_neg}
\end{figure}

@@

Our feature vectors have been designed based on insights in the automated face detection literature \citep{ViolaJones}.  The basic strategy is to encode a very large number ($\sim$40,000) of generic statistics about extended structures at various positions and scales. This strategy seems to be more effective than trying to tune a feature vector to the specific object being identified. While most elements in the feature vector will have no predictive power in the classification, the appropriate learning algorithm will simply ignore this information.

@@

\begin{enumerate}
\item The wavelet coefficients from the Discrete Cosine and Daub4 wavelet transforms. These coefficients encode how much power is in various spatial frequencies.
\item The image dot product\footnote{i.e., the sum of the pixel-by-pixel product of two images} of each image with 49 template images of circular rings of different size and thickness. Bubbles are morphologically similar to these templates, and presumably have larger dot products than a random intensity distribution.
\item The byte size of each compressed postage stamp image -- images substantially more or less smooth than a bubble compress to smaller and larger files, respectively.
\item The DAISY features of the image \citep{DAISY}. DAISY features are derived from measurements of local intensity gradients, and encode information about simple shapes in images; they are conceptually similar to edge detection filters.
\end{enumerate}

\subsection{Training}
While there are many algorithms designed to learn classifications form a training set of feature vectors, we focus here on the Random Forest algorithm \citep{Breiman01}. Random Forests are aggregates of a simpler learning algorithm called a decision tree. A decision tree is conceptually similar to the game twenty questions: in this game, one player asks a series of yes/no questions in an attempt to deduce the identity of a hidden object selected by another player. Likewise, a decision tree classifies an object by performing several boolean tests on elements of that object's feature vector (e.g., $F_3 > 5$). These tests are arranged in a binary tree. The decision tree algorithm traverses this tree for each object, traversing the left or right subtree depending on whether each boolean test is True or False. The leaves of the tree contain the final classification (in the present context, ``bubble'' or ``not bubble'').

The decision tree is constructed from the training data, where the feature vectors and classification are both known. Trees are built one node at a time, in a greedy fashion -- that is, at every step of the learning process, a node is added to the tree to maximize the tree's classification performance on the training data. Several different heuristics can be used to measure classification performance.

On their own, decision trees are prone to over-fitting the training data by adding too many nodes. In a more traditional fitting context, this is analogous to adding too many terms to a polynomial fit of noisy data -- in both situations, the model describes the input data very well, but does not generalize to new predictions. Random Forests overcome this problem by building several trees on different subsets of the data. Since over-fitting is sensitive to the exact input data used, each individual tree in a Random Forest over-fits in a different way. By averaging the classifications of each tree, the Random Forest mitigates over-fitting. Random Forests have proven effective in many machine learning contexts \citep{Kuhn13}.

In addition to good performance and robustness against over fitting, the Random Forest algorithm has several other advantageous properties: it is possible to interpret how important each element in the feature vector is to the classification, the algorithm naturally ignores irrelevant elements of the feature vector, and it is conceptually and computationally simple.

\subsection{Hyperparameter Optimization}
The Random Forest algorithm has a small number of tunable parameters. The most important of these are the number of trees in the forest, and the heuristic used to measure classification quality when building each tree. We determined good settings for these ``hyper-parameters'' using cross validation. During cross validation, the training data is partitioned into two sets. The model is fit to the first set using a given set of hyper-parameters, and is evaluated on the second set. This process repeats for each combination of hyper-parameters, and the combination with the best performance on the test data is used for the final model. From our cross-validation analysis, we chose forests with 800 trees, using the information gain heuristic \citep{Raileanu04}.

@@

Classifiers cannot be used to re-classify the data used to train them, due to the possibility of over-fitting. Thus, to obtain valid classifications for all fields, we constructed three classifiers. Each classifier was trained using data from two-thirds of the sky, and used to classify the remaining third. The three training zones are interleaved longitude sections, so that each classifier is sensitive to trends with longitude. Figure \ref{fig:zone} depicts how the $0 < \ell < 10$ region is partitioned -- the shaded regions denote the portion of the sky used to train each classifier. The three classifiers are identical apart from the longitude zones used for training and classification.

\begin{figure}[h!]
\includegraphics{zones}
\caption{Illustration of the zones used to train each classifier. Each classifier is trained using the data from the shaded regions in its zone, and used to classify the light regions.}
\label{fig:zone}
\end{figure}

@@

One standard way to summarize the performance of a classifier is to plot the false positive rate (fraction of negative examples incorrectly classified) versus the true positive rate (fraction of bubbles correctly classified) as the threshold varies. This is called the Receiver Operating Characteristic, and is shown for the three classifiers in Figure \ref{fig:roc}. The false positive rate is measured by classifying $\sim 15000$ random negative fields, and the true positive rate is measured using the high-quality subset of the MWP catalog described above. Recall that the set of bubbles in this test set were selected because they are very well defined. Thus, Figure \ref{fig:roc} overestimates the true positive rate when presented with more ambiguous bubbles. We compare the performance of the classifier on more representative bubbles in Section \ref{sec:expert}.

\begin{figure}[h!]
\includegraphics{roc}
\caption{The tradeoff between false positive rate and true positive rate, for the three classifiers in Brut. Each classifier is trained using a different zone, depicted in Figure \ref{fig:zone}.}
\label{fig:roc}
\end{figure}

@@

Using our automatic bubble classifier to conduct a full search for bubbles in \textit{Spitzer} data is straightforward. The main task involves scanning the classifier over all images, at all scales. At each of these regions, the classifier yields a score, representing its confidence that the field contains a bubble. A simple threshold on this score identifies regions likely to contain a bubble -- from the discussion of Figure \ref{fig:roc} above, adjusting this threshold sets the tradeoff between completeness and reliability. Note that Brut only attempts to determine the location and size of bubbles. Citizen scientists, on the other hand, were able to constrain the aspect ratio, position angle, and thickness of each object.

In practice, the classifier is insensitive to small adjustments to the size or location of the field. As a result, bubbles are usually associated with a cluster of regions that rise above the detection threshold (the white circles in Figure \ref{fig:cluster}). We follow a simple procedure to merge these individual detections. Given a collection of raw detections, this procedure identifies the two nearest locations at similar scales, and discards the location assigned a lower score by the classifier. This process repeats until no two fields are close enough to discard. Two fields are close enough to merge if they are separated by less than the radius of either field, and have sizes that differ by less than 50\%. The red circle in Figure \ref{fig:cluster} shows the result of merging detections.

\begin{figure}[h!]
\includegraphics{cluster}
\caption{An illustration of how Brut merges multiple bubble detections. Each white circle is a location with a high Brut score. The red circle is the region with the highest brut score, and is used as the final detection.}
\label{fig:cluster}
\end{figure}

@@
